Blaming the Typeless: \\ Scalable, Human-Centric Python Localization

### 2018-08-21 paper/main.tex
Wes pass over introduction

\emph{Placeholder Outline.} This text is an informal placeholder outline
only.

\emph{Identify an important problem.} Support the claim that Python fault
localization for possibly-ill-typed student fragments is an important
problem. Likely steps: fault localization is important, Python is
important, incomplete code is important. First, software engineering is
important and expensive; testing, maintenance, and debugging are the
dominant activities in SE; localization and triage are key steps. Second,
Python is increasingly used for pedagogy and industrial deployment, here
are some GitHub numbers about the rate of growth in Python projects
compared to C or Java, etc. Third, IDE tool use requires operating on code
fragments, here are some numbers from PythonTutor~\cite{Guo2013-vu}, generality, etc.

\emph{Identify the properties of a good solution.} Pick your favorite
three. Some candidates: accuracy (cite Parnin/Orso paper an indicate that
it must be found within top three, etc., since humans will not read long
lists), generality (must apply to student code and incomplete fragments),
scalability (must apply to hundreds of thousands of fragments, must present
real-time responses suitable for web/IDE use), agreement with subjective
judgments (must handle ``ambiguity'' the same way humans would when you
don't know whether to blame the use or the definition, etc.), efficiency
(must operate using a small number of dynamic runs, unlike Tarantula,
etc.).

\emph{Show that the current state of the art is inadequate.} For each $x$ of
\{manual debugging, Python's default error messages,
Mycroft/Sherrloc/Tarantula/whatever\}, show that $x$ lacks at least one of the
desired properties listed above.

\emph{Insights and special sauce.} What are the two or three good ideas
that we bring together to solve this problem? Some candidates: the use of
modern machine learning techniques (deep learning, etc.); the use of
tree-based contextual features to hit the sweet spot between syntax and semantics;
special handling of unsatisfied use-def type constraints (TBA); etc.

\emph{What is our proposal?} We propose XYZ, an algorithm to do ABC. It
takes as input DEF. It has stages P, Q and R. It produces as output GHI. It
provides guarantees JKL.

\emph{How will we evaluate it?} For each of the properties of a good
solution listed above, indicate how we will evaluate our contribution. What
evidence will we provide? Typically, for each property $x$, we support our
claim to $x$ with either a formal proof (e.g., our algorithm has $x$ by
construction), an empirical evaluation on software artifacts, or an
empirical evaluation on human subjects. In each case, indicate both the
evaluation metrics (e.g., if we care about efficiency, is it measured in
lines-per-second? if we care about precision, what is our non-standard
top-three metric) and success criteria (e.g., how will we define success?
what is the baseline we want to beat). For empirical evaluations, briefly
and implicitly argue that the results are likely to generalize by giving
the size of the sampled set.

\emph{Introduction wrap-up.} The contributions of this paper are as follows
...

>>>>

Dynamically-typed languages are becoming increasingly common for rapid
application development, full-scale software engineering, and
pedagogy (e.g.,~\cite{FIXME,FIXME}). Languages like Python, Ruby and Lua
often feature a gentle learning curve and attractive facilities, from
familiar garbage collection and regular expressions to more expressive and
exotic list comprehensions and higher-order function blocks. One side
effect of popularity with novices and students, however, is that many
programmers of such languages may lack expertise when presented with
dynamic type errors, and may thus struggle to interpret and localize
them~\cite{FIXME,FIXME}. We propose a machine learning approach based on
static, dynamic and contextual information that effectively localizes
beginner type errors.

The relative costs and benefits of static vs. dynamic error detection are
well-studied. Despite this, however, many currently-available techniques
are ill-suited to address the problem of localization for type errors in
dynamic languages. First, relying on the extant language interpreter
requires test inputs~\cite{FIXME}, may involve error messages that are
difficult for novices to understand~\cite{FIXME}, and may misleadingly
implicate the symptom rather than the cause~\cite{FIXME}.  Second, a
significant amount of development is carried out via IDE or web interfaces,
which require tools that operate on code fragments~\cite{Guo2013-vu}.
Third, attempts to retrofit static type systems into dynamic languages,
such as TypeScript for JavaScript~\cite{FIXME} and FIXME for
Python~\cite{FIXME} are promising but not yet as widely
deployed~\cite{FIXME}. Finally, research tools and algorithms for fault
localization either require rich static types (e.g., as in
Mycroft~\cite{FIXME} or Sherrloc~\cite{FIXME}), require test cases (e.g.,
as in Tarantula~\cite{FIXME}), or produce voluminous ranked output lists
that have been found to be unhelpful in general and less useful to novices
in particular~\cite[Sec.~5.1]{orso-parnin}.

We focus on Python as an indicative, popular dynamically-typed language.
We desire a fault localization algorithm for Python type errors that
will agree with human intuition (\emph{accuracy}), will apply to
off-the-shelf, unannotated beginner-written program fragments
(\emph{generality}), and will scalably apply to industrial-strength
deployments (\emph{scalability}). Our key insights are (1) that we can leverage
modern supervised machine learning for accuracy and an agreement with human
norms; (2) that we can use static, dynamic, contextual and slice-based
features to generalize to off-the-shelf beginner programs; and (3) that we
can use large beginner datasets, such as those from
PythonTutor.com~\cite{Guo2012-vu}, to assess scalability.

We propose an algorithm that takes an input a Python program fragment that 
produces an uncaught runtime exception such as a type error. Via machine
learning over a carefully-selected set of program-relevant features, we
produce the expression most likely implicated in the exception according to
our learned model. We evaluate on an entire year of buggy Python 3 program
executions on PythonTutor.com --- over 270,000 instances --- measuring
accuracy with respect to the human actions as the ground truth and
comparing against the vanilla Python interpreter as a baseline.

The contributions of this paper are as follows:
\begin{enumerate}

\item We present an algorithm for accurately localizing dynamic type errors
in Python. We find that FIXME.

\item We describe and evaluate static, dynamic, contextual and slice-based
features for generalizing machine-learning fault localization to
beginner-written Python fragments. We find that FIXME.

\item We analyze the results of a thorough evaluation on over 270,000 real
user instances, demonstrating the scalability of our algorithm.
